Stateless Model Checking with Data-Race Preemption Points

### 2015-10-17 pldi-16/di.tex
write about free-re-malloc and add figure 1

\section{Design}

@@

% define minimal state space and maximal state space

\subsection{Filtering false positive data races}

{\bf Terminology.}
Importantly, we distinguish data-race {\em candidates} and data-race {\em bugs}.
Because data-race analysis is false-positive-prone, Landslide reports unprotected access pairs separately from the problems listed above, calling the pair a {\em data race candidate}.
Should a future interleaving, preempting during those accesses, cause a failure (e.g. assertion or deadlock), then we report a {\em data-race bug}.
Otherwise, if the access pair can be reordered, but does not produce a failure under any interleaving, it is a {\em benign data race}.
If they cannot be reordered at all, it is a {\em false positive}.
For brevity, we classify benign data races as a subset of false positives.

@@

We chose \landslide~\cite{landslide} as our stateless model checker due to its ability to trace program execution at the granularity of individual instructions and memory accesses, which is mandatory for data-race analysis~\cite{eraser}.
% optional: cite tsan above
\landslide~features Dynamic Partial Order Reduction (DPOR) \cite{dpor}, state space estimation \cite{estimation}, and data-race detection.
It avoids state space cycles (e.g. ad-hoc synchronization with {\tt yield()} or even {\tt xchg} loops) similarly to Fair-Bounded Search \cite{bpor}.
% this line can be cut if space is needed
It can test both user- and kernel-level code, though the latter requires user annotations and is limited to timer-driven nondeterminism.
Its bug-detection metrics include assertion failures, deadlock, segfaults, heap checking (like Valgrind~\cite{valgrind}), and a heuristic infinite loop/livelock checker.

@@

When \quicksand~receives a data race report, it adds two new jobs to its workqueue:
a ``small'' job to preempt on the racing instruction only,
and a ``big'' job to preempt on that instruction as well as each PP used by the reporting job.
%
Hence each {\em pair} of racing accesses will spawn four new jobs, as shown in Figure~\ref{fig:new-dr-jobs}.
Rather than adding a maximal job with both new PPs at once, we prefer to add smaller jobs which have a higher chance of completing in time.
If those state spaces are bug-free, they will in turn add the maximal job later.
%
The rationale of spawning multiple jobs is that which will be more fruitful cannot be known in advance:
while the big job risks not completing in time,
the small job risks missing the data race entirely if the original PPs were required to expose it.
In practice, we observed some bugs found quickly by these small jobs, and other bugs missed by the small jobs found eventually by the big jobs.
% which justifies this design

>>>>

% TODO: get latex to put this on page 2 somehow
\begin{figure*}[t]
	\includegraphics[width=0.15\textwidth]{tree0.pdf}
	\includegraphics[width=0.225\textwidth]{tree1.pdf}
	\includegraphics[width=0.25\textwidth]{tree2.pdf}
	\includegraphics[width=0.35\textwidth]{tree3.pdf}
	\caption{Iterative Deepening visualization.
		The minimal state space includes only mandatory thread switches, such as {\tt yield()} or {\tt cond\_wait()}~\cite{landslide}.
Multiple further tests can be run: preempting on calls to {\tt mutex\_lock} alone, {\tt mutex\_unlock} alone, or both together.
Each option increases the state space size by an unpredictable factor, so multiple state spaces should be tested in parallel.
Estimation techniques~\cite{estimation} inform which state spaces to prioritize.
}
	% TODO: reference this label in first few paragraphs of design
	\label{fig:id}
\end{figure*}
\section{Design}

@@

% define minimal state space and maximal state space
{\bf Terminology.}
Importantly, we distinguish data-race {\em candidates} and data-race {\em bugs}.
Because data-race analysis is false-positive-prone, Landslide reports unprotected access pairs separately from the problems listed above, calling the pair a {\em data race candidate}.
Should a future interleaving, preempting during those accesses, cause a failure (e.g. assertion or deadlock), then we report a {\em data-race bug}.
Otherwise, if the access pair can be reordered, but does not produce a failure under any interleaving, it is a {\em benign data race}.
If they cannot be reordered at all, it is a {\em false positive}.
For brevity, we classify benign data races as a subset of false positives.

\subsection{Verifying data races as buggy or benign}

\subsection{Suppressing malloc-related false positives}

% TODO: Understand what prior work means by 'happens before'. It's different from our MUST happen before relation.
% TODO: Understand how it relates to this. Would it scoop us? Or are we outside their model?

We identify a particular class of false positive data-race candidate in which the associated memory is recycled by {\tt malloc} between the two accesses.
Figure~\ref{fig:free-re-malloc} shows a common code pattern and interleaving which can expose such behavior.
If the {\tt malloc} on line 4 returns the same address passed to {\tt free} on line 2, then lines 1 and 6 will be flagged as a data race.
To the human eye, this is obviously a false positive: reordering lines 4-6 before lines 1-2 will change {\tt malloc}'s return value, causing {\tt x} and {\tt y} to no longer collide.
Here, Thread 2's logic usually corresponds to the initialization pattern \cite{eraser}, but for generality we have added a {\tt publish} action on line 4.

\begin{figure}[t]
\begin{tabular}{rll}
	& \multicolumn{2}{c}{\texttt{struct x \{ int foo; int baz; \};}} \\
	& \multicolumn{2}{c}{\texttt{struct y \{ int bar; \};~~~~~~~~~~}} \\
	\\
	& Thread 1 & Thread 2 \\
	1 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	2 & \texttt{\hilight{olivegreen}{free}(x);} \\
	3 & & \texttt{// x's memory recycled} \\
	4 & & \texttt{y~=~\hilight{olivegreen}{malloc}(...);} \\
	5 & & \texttt{publish(y);} \\
	6 & & \texttt{\hilight{brickred}{y->bar = ...;}} \\
\end{tabular}
\caption{A common pattern with {\tt malloc()} that produces false positive data race candidates.}
\label{fig:free-re-malloc}
\end{figure}

% This class of false positive is unique to heap-allocated memory, among all ways threads could communicate. By contrast, global memory has unlimited lifetime, and message-passing primitives enforce a must-happens-before relationship which precludes the race.

It is quite simple to mechanically recognize when {\tt x} and {\tt y} correspond to different abstract allocations despite colliding on address.
\landslide~implements this check by adding a generation counter to its heap state tracking:
each allocation is given a unique ID, and when evaluating whether two heap accesses can race, their generation counters must match.

However, when limited to a single execution, suppressing any data race matching this pattern is unsound.
Consider the more unusual program in Figure~\ref{fig:frm-bug}:
Now, the memory is recycled the same way, but the racing access's address is not tied to {\tt malloc}'s return value.
Hence, reordering lines 6-7 before line 3 will cause {\tt x} and {\tt x2} to race.

\begin{figure}[t]
\begin{tabular}{rll}
	& Thread 1 & Thread 2 \\
	1 & \texttt{publish(x);} & \\
	2 & & \texttt{x2 = get\_published\_x();} \\
	3 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	4 & \texttt{\hilight{olivegreen}{free}(x);} \\
	5 & & \texttt{// x's memory recycled} \\
	6 & & \texttt{y~=~\hilight{olivegreen}{malloc}(...);} \\
	7 & & \texttt{\hilight{brickred}{x2->foo = ...;}} \\
\end{tabular}
\caption{blah blah}
\label{fig:frm-bug}
\end{figure}

% TODO: Figure out why they claim "Happens before produces NO false positives, only benign races".
% It seems impossible.
% But if true, it means either (a) they can somehow identify FRM DRs on the 1st pass, not needing to replay,
% or (b) reuse of memory by malloc is somehow outside their concurrency model.
Note that while replay-oriented data race tools can already classify free-re-malloc candidates as false positives % benign? (not benign. they are ghosts. not even real.)
\cite{recordreplaydrs},
they require replaying the threads in a new interleaving.
Moreover, \cite{portend} argues that accurate classification may require re-executing many times,
according to many pre- and post-race sequences,
which is tantamount to adding a new state space in \quicksand.
Our goal here is to eliminate free-re-malloc false positives immediately upon detecting them, with no replay required.

@@

We chose \landslide~\cite{landslide} as our stateless model checker due to its ability to trace program execution at the granularity of individual instructions and memory accesses, which is mandatory for data-race analysis~\cite{eraser}.
% optional: cite tsan above
\landslide~features Dynamic Partial Order Reduction (DPOR) \cite{dpor}, state space estimation \cite{estimation}, and data-race detection.
It avoids state space cycles (e.g. ad-hoc synchronization with {\tt yield()} or even {\tt xchg} loops) similarly to Fair-Bounded Search \cite{bpor}. % TODO: devote a paragraph to this, after talking about drs in locks
% this line can be cut if space is needed
It can test both user- and kernel-level code, though the latter requires user annotations and is limited to timer-driven nondeterminism.
Its bug-detection metrics include assertion failures, deadlock, segfaults, heap checking (like Valgrind~\cite{valgrind}), and a heuristic infinite loop/livelock checker.

@@

When \quicksand~receives a data race report, it adds two new jobs to its workqueue:
a ``small'' job to preempt on the racing instruction only,
and a ``big'' job to preempt on that instruction as well as each PP used by the reporting job.
%
Hence each {\em pair} of racing accesses will spawn four new jobs, as shown in Figure~\ref{fig:new-dr-jobs}.
Rather than adding a maximal job with both new PPs at once, we prefer to add smaller jobs which have a higher chance of completing in time.
If those state spaces are bug-free, they will in turn add the maximal job later.
%
The rationale of spawning multiple jobs is that which will be more fruitful cannot be known in advance:
while the big job risks not completing in time,
the small job risks missing the data race entirely if the original PPs were required to expose it.
In practice, we observed some bugs found quickly by these small jobs, and other bugs missed by the small jobs found eventually by the big jobs,
which motivates the need for Iterative Deepening to decide at runtime which to prioritize.
% TODO: Put numbers here.
% which justifies this design

### 2015-10-17 pldi-16/evaluation.tex
write about free-re-malloc and add figure 1

Although \quicksand~presents Iterative Deepening and data-race PPs as interconnected techniques, they each could theoretically be employed alone in other model checkers.
For example, a single-state-space tool could use data-race candidates during immediately subsequent interleavings, essentially changing the state space on the fly.
Likewise, a message-passing-only tool could employ Iterative Deepening despite data races being absent from its concurrency model.
Hence, though many of our experiments compare \quicksand~to the state-of-the-art as a whole,
we also sought to evaluate each technique individually.
Our evaluation answers the following questions:
\begin{enumerate}
	\item Does Iterative Deepening find bugs faster than SSS-MC in subset state spaces, even without data-race PPs?
			% Probably not... ICB is state of the art here.
			% In large tests, can Iterative Deepening provide partial verification by completing smaller state spaces
	\item Is combining data-race analysis with MC productive?
		\begin{enumerate}
			\item Do data-race PPs expose new bugs that couldn't be found with SSS-MC's fixed-PP-set approach?
				% Elaborate later:
				% Among those, how many were missed in a {\em completed} execution of the otherwise ``maximal'' state space?
			\item Do we avoid false positives compared to a single-execution data race analysis?
				% TODO: Argue:
				% It is fair to compare multiple pass DR-analysis under Landslide against just a single execution because prior work DR detectors, being not integrated with a MC, are not intended to uncover different results under subsequent runs.
				% Define a "stress tester" as a class of bug detectors where they are intended to [...]
				% Maybe say: Do there exist any stress-tester DR detectors where they are intended to produce new results under reruns?

				% Explain later as:
				% How many data-race candidates were verified as benign
				% But to be fair, you have to count how many DRs are reported as "couldn't test these, check yourself" at the end.
				% Also Include:
				% How many false positives does the free-re-malloc technique suppress?
				% TODO: If you have time, re-run all of the dr-only bug tests, with DR_FALSE_NEG enabled, and see how much fewer bugs get found (how many bugs get pushed past the time limit?)
			% TODO: This one's optional. You can give up on it.
			\item Do we find data-race bugs that would be false-negatives during a single-execution analysis?%Do we avoid false negatives compared to single-pass?
				%TODO: for this experiment, set EXPLORE_BACKWARDS=0
				% TODO: And disable false-neg malloc-free technique
				% TODO: And also disable the confirmed/suspected thing
				% (where mem.c waits for reorder observed before
				% messaging the latter half of the DR to QS)
		\end{enumerate}
\end{enumerate}
\subsection{Test Suite}
% TODO: Maybe say how many lines of code total? How many lines on average per P2/pintos? (careful with pintos; lots of basecode)
Our test suite consists of \numthrlibs~``P2'' student thread libraries, from CMU's 15-410 OS class,
%across the Spring and Fall 2014 and Spring 2015 semesters;
and \numpintoses~``Pintos'' student kernels, from Berkeley's CS162 and U. Chicago's CS230 OS classes.
%
The P2 thread library comprises \texttt{thr\_create()}, \texttt{thr\_exit()}, \texttt{thr\_join()}, mutexes, condvars, semaphores, and r/w locks;
all implemented from scratch in userspace with a UNIX-like system call interface \cite{kspec,thrlib}.
%
The Pintos kernel project
involves implementing priority scheduling, \texttt{sleep()}, and user-space process management (\texttt{wait()} and \texttt{exit()})
using provided bare-bones mutex, context-switch, and virtal memory implementations
\cite{pintos}.
% P2 SLOC stats: 1807 avg; 1723 median; range 1181-4114.
% All numbers, obtained with:
% cd p2s; for i in */*; do wc -l $i/user/libthread/*.{c,h,S} $i/user/libthread/*/*.{c,h,S} ; done | grep total
% 1181 1192 1221 1230 1238 1240 1243 1261 1275 1307 1310 1318 1325 1334 1336 1345 1366
% 1388 1388 1403 1415 1416 1430 1451 1478 1498 1527 1589 1618 1635 1638 1654 1675 1676
% 1716 1719 1720 1723 1723 1727 1737 1743 1744 1751 1769 1777 1782 1789 1789 1812 1918
% 1926 1946 1994 2022 2043 2066 2077 2088 2099 2131 2164 2172 2190 2215 2227 2277 2282
% 2384 2387 2483 2486 2503 2514 2551 2597 2610 2665 4114
Though not ``real world'' programs, both projects are quite large: % maybe "complex"?
the P2s average 1807 lines of C and x86 assembly (stddev 489.5),
% Pintos SLOC stats: TODO
and the Pintoses average {\bf 9999999} % TODO

\newcommand\mxtest{\texttt{mx\_test}}
\newcommand\tej{\texttt{thr\_exit\_join}}
\newcommand\bct{\texttt{broadcast}}
\newcommand\paraguay{\texttt{paraguay}}
\newcommand\paradise{\texttt{paradise\_lost}}
\newcommand\rwl{\texttt{rwl\_test}}
We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness, \tej, a test of thread lifecycle, \bct~and \paraguay, for condvars, \paradise~for semaphores, and \rwl~for r/w locks.
For \mxtest~we enabled \landslide's mutex-testing option described in \sect{\ref{sec:landslide}}.
%, which changes the lock-set tracking to consider accesses within {\tt mutex\_lock} and {\tt mutex\_unlock} as data-race candidates.
%(Normally, they are considered protected by the very lock they implement: assuming the locks are already correct to enable productive data-race analysis on the rest of the codebase.)
\newcommand\prisema{\texttt{priority\_sema}}
\newcommand\waitsimple{\texttt{wait\_simple}}
% TODO: Add more test cases
We tested the Pintoses with 2 programs from the class test suite: \prisema, a test of the kernel scheduling algorithm, and \waitsimple, a test of process lifecycle. % TODO talk more about these.
In total, this test suite comprises {\large \bf 99999} % TODO
unique state spaces.

All tests were run on 12-core 3.2 GHz Xeon machines with 12GB of RAM.
\quicksand~tests were given 10 CPUs for 1 hour each.
To compare to SSS-MC, we ran a control experiment for each test, running \landslide~on a single state space with all PPs enabled in advance (and no data-race PPs).
% TODO figure out somewhere to mention what landslide's pps are: mx lock/unlock (aka sema up/down)
Though \landslide~does not implement parallel DPOR \cite{parallel-dpor}, we leveled the playing field by giving each control 1 CPU for 10 hours and dividing all associated times by 10 (simulating perfect parallelism).

\begin{table}[t]
	\begin{tabular}{l|l|l}
			& QS bugs & SSS-MC bugs \\
		\hline
		\mxtest & eg 1000 & eg 0 \\
		\bct & & \\
		etc... & & \\
		\hline
		Total & & \\
	\end{tabular}
	\caption{Comparison of all bugs found, broken down by test case, among all P2s (top 6) and Pintoses (bottom {\bf \large 9999999})}
	\label{tab:allbugs}
\end{table}

@@

asdfadsfasdf

\subsection{Evaluating Data-Race PPs}

In this section we show that using data-race PPs with \landslide~is more effective than either SSS-MC or single-pass data-race detection alone.

{\bf Finding bugs missed by SSS-MC.}
Though state-of-the-art MCs preempt only on synchronization events, many serious concurrency bugs are caused by data races leading to corrupted shared state.
In Table~\ref{tab:drbugs} we count ...... % TODO

@@

{\bf Avoiding false positive data-race candidates.}
% Though we mechanically verify whether each data race candidate leads to a bug, each new PP can increase combinatorially..... obviously wish to avoid...

{\bf Avoiding false negative data-race bugs.}
\label{sec:eval-falseneg}
Some pairs of unprotected shared-memory accesses may be hidden in some control flow path that requires a nondeterministic preemption to be executed.
In such cases, a single-pass data-race detector
%could not achieve the coverage necessary
would fail
to identify the access pair as a data-race candidate to begin with.
%
We counted how many such data-races, used as PPs, would lead to \landslide~finding new bugs,
thereby making them {\em false negatives} of the single-pass approach.
% TODO: Put a figure here giving an example of where e.g. a data race only shows up during the contention path of a mutex.
We classified each data-race candidate according to whether \landslide~had reported them during the first interleaving, before any backtracking or preempting: if so, they were {\em deterministic data races} (hence could be found by single-pass).
%
To ensure a fair comparison, we disabled \landslide's {\em false-positive}-avoidance techniques during this experiment.
For example, we reported free-re-malloc data races during the first interleaving as {\em deterministic}, as a single-pass analysis must, rather than waiting until future interleavings to confirm them (see \sect{\ref{sec:free-re-malloc}}).

@@

% TODO: Future work.
% Future work: Add a way to configure even smaller subsets (eg "only mutex_locks called from site X") for cases where mx_lock and mx_unlock alone are still too big. Count the number of kernels for which this was the case.

>>>>

Although \quicksand~presents Iterative Deepening and data-race PPs as interconnected techniques, they each could theoretically be employed alone in other model checkers.
For example, a single-state-space tool could use data-race candidates during immediately subsequent interleavings, essentially changing the state space on the fly.
Likewise, a message-passing-only tool could employ Iterative Deepening despite data races being absent from its concurrency model.
Hence, though many of our experiments compare \quicksand~to the state-of-the-art as a whole,
we also sought to evaluate each technique individually.
Our evaluation answers the following questions:
\begin{enumerate}
	\item Does Iterative Deepening find bugs faster than SSS-MC in subset state spaces, even without data-race PPs?
			% Probably not... ICB is state of the art here.
			% In large tests, can Iterative Deepening provide partial verification by completing smaller state spaces
	\item Is combining data-race analysis with MC productive?
		\begin{enumerate}
			\item Do data-race PPs expose new bugs that couldn't be found with SSS-MC's fixed-PP-set approach?
				% Elaborate later:
				% Among those, how many were missed in a {\em completed} execution of the otherwise ``maximal'' state space?
			\item Do we avoid false positives compared to a single-execution data race analysis?
				% Explain later as:
				% How many data-race candidates were verified as benign
				% But to be fair, you have to count how many DRs are reported as "couldn't test these, check yourself" at the end.
				% Also Include:
				% How many false positives does the free-re-malloc technique suppress?
				% TODO: If you have time, re-run all of the dr-only bug tests, with DR_FALSE_NEG enabled, and see how much fewer bugs get found (how many bugs get pushed past the time limit?)
			% TODO: This one's optional. You can give up on it.
			\item Do we find data-race bugs that would be false-negatives during a single-execution analysis?%Do we avoid false negatives compared to single-pass?
				%TODO: for this experiment, set EXPLORE_BACKWARDS=0
				% TODO: And disable false-neg malloc-free technique
				% TODO: And also disable the confirmed/suspected thing
				% (where mem.c waits for reorder observed before
				% messaging the latter half of the DR to QS)
		\end{enumerate}
\end{enumerate}
\subsection{Test Suite}
% TODO: Maybe say how many lines of code total? How many lines on average per P2/pintos? (careful with pintos; lots of basecode)
Our test suite consists of \numthrlibs~``P2'' student thread libraries, from CMU's 15-410 OS class,
%across the Spring and Fall 2014 and Spring 2015 semesters;
and \numpintoses~``Pintos'' student kernels, from Berkeley's CS162 and U. Chicago's CS230 OS classes.
%
The P2 thread library comprises \texttt{thr\_create()}, \texttt{thr\_exit()}, \texttt{thr\_join()}, mutexes, condvars, semaphores, and r/w locks;
all implemented from scratch in userspace with a UNIX-like system call interface \cite{kspec,thrlib}.
%
The Pintos kernel project
involves implementing priority scheduling, \texttt{sleep()}, and user-space process management (\texttt{wait()} and \texttt{exit()})
using provided bare-bones mutex, context-switch, and virtal memory implementations
\cite{pintos}.
% P2 SLOC stats: 1807 avg; 1723 median; range 1181-4114.
% All numbers, obtained with:
% cd p2s; for i in */*; do wc -l $i/user/libthread/*.{c,h,S} $i/user/libthread/*/*.{c,h,S} ; done | grep total
% 1181 1192 1221 1230 1238 1240 1243 1261 1275 1307 1310 1318 1325 1334 1336 1345 1366
% 1388 1388 1403 1415 1416 1430 1451 1478 1498 1527 1589 1618 1635 1638 1654 1675 1676
% 1716 1719 1720 1723 1723 1727 1737 1743 1744 1751 1769 1777 1782 1789 1789 1812 1918
% 1926 1946 1994 2022 2043 2066 2077 2088 2099 2131 2164 2172 2190 2215 2227 2277 2282
% 2384 2387 2483 2486 2503 2514 2551 2597 2610 2665 4114
Though not ``real world'' programs, both projects are quite large: % maybe "complex"?
the P2s average 1807 lines of C and x86 assembly (stddev 489.5),
% Pintos SLOC stats: TODO
and the Pintoses average {\bf 9999999} % TODO

\newcommand\mxtest{\texttt{mx\_test}}
\newcommand\tej{\texttt{thr\_exit\_join}}
\newcommand\bct{\texttt{broadcast}}
\newcommand\paraguay{\texttt{paraguay}}
\newcommand\paradise{\texttt{paradise\_lost}}
\newcommand\rwl{\texttt{rwl\_test}}
We tested P2s with 6 multithreaded programs:
% from the 410 test suite % XXX: I would like to say this but this is a lie; figure out what else i can say instead
% each tailored to exercise a different part of the P2 project
\mxtest, for locking algorithm correctness, \tej, a test of thread lifecycle, \bct~and \paraguay, for condvars, \paradise~for semaphores, and \rwl~for r/w locks.
For \mxtest~we enabled \landslide's mutex-testing option described in \sect{\ref{sec:landslide}}.
%, which changes the lock-set tracking to consider accesses within {\tt mutex\_lock} and {\tt mutex\_unlock} as data-race candidates.
%(Normally, they are considered protected by the very lock they implement: assuming the locks are already correct to enable productive data-race analysis on the rest of the codebase.)
\newcommand\prisema{\texttt{priority\_sema}}
\newcommand\waitsimple{\texttt{wait\_simple}}
% TODO: Add more test cases
We tested the Pintoses with 2 programs from the class test suite: \prisema, a test of the kernel scheduling algorithm, and \waitsimple, a test of process lifecycle system calls.
In total, this test suite comprises 632 unique state spaces.

All tests were run on 12-core 3.2 GHz Xeon machines with 12GB of RAM.
\quicksand~tests were given 10 CPUs for 1 hour each.
To compare to SSS-MC, we ran a control experiment for each test, running \landslide~on a single state space with all PPs on sync primitives enabled in advance (and no data-race PPs).
% TODO figure out somewhere to mention what landslide's pps are: mx lock/unlock (aka sema up/down)
Though \landslide~does not implement parallel DPOR \cite{parallel-dpor}, we compensated by giving each control test 1 CPU for 10 hours, then dividing all associated times by 10 (simulating perfect parallelism).

\begin{table}[t]
	\begin{tabular}{l|l|l}
			& QS bugs & SSS-MC bugs \\
		\hline
		\mxtest & eg 1000 & eg 0 \\
		\bct & & \\
		etc... & & \\
		\hline
		Total & & \\
	\end{tabular}
	\caption{Comparison of all bugs found, broken down by test case, among all P2s (top 6) and Pintoses (bottom 2)}
	\label{tab:allbugs}
\end{table}

@@

% TODO: If you have time, rerun all the quicksand experiments JUST running the 4 base state spaces. Give it 10/4 hours of cpu budget on 4 cores.

First we compare \quicksand~to SSS-MC
% mention exactly which state spaces we are comparing here

% mention partial verification in terms of state space completion, when SSSMC times out.

\subsection{Finding data-race bugs missed by SSS-MC}

%In this section we show that using data-race PPs with \landslide~is more effective than either SSS-MC or single-pass data-race detection alone.

Though state-of-the-art MCs preempt only on synchronization events, many serious concurrency bugs are caused by data races leading to corrupted shared state.
Hence, in Table~\ref{tab:drbugs} we count how often \quicksand~uncovered a bug only in state spaces which included data-race PPs,
while

@@

\subsection{Avoiding false positive data-race candidates}
% Though we mechanically verify whether each data race candidate leads to a bug, each new PP can increase combinatorially..... obviously wish to avoid...

We also counted the number of free-re-malloc false positives that \landslide~suppressed.

\subsection{Finding nondeterministic data-race candidates}
\label{sec:eval-falseneg}
Some memory accesses may be hidden in a control flow path that requires a nondeterministic preemption to be executed.
In such cases, a single-pass dynamic data-race detector
%could not achieve the coverage necessary
would fail
to identify a racing access pair as a candidate to begin with.
%
We counted how many such data-races, used as PPs, led to \quicksand~finding new bugs,
thereby making them {\em false negatives} of the single-pass approach.
% TODO: Put a figure here giving an example of where e.g. a data race only shows up during the contention path of a mutex.
We classified each data-race candidate according to whether \landslide~had reported them during the first interleaving, before any backtracking or preempting: if so, they were {\em deterministic data races} (hence could be found by single-pass).

To ensure a fair comparison, we disabled \landslide's {\em false-positive}-avoidance techniques during this experiment.
For example, we reported free-re-malloc data races during the first interleaving as {\em deterministic}, as a single-pass analysis must,
rather than waiting until future interleavings to confirm them (as explained in \sect{\ref{sec:free-re-malloc}}).

				% TODO: Argue:
				% It is fair to compare multiple pass DR-analysis under Landslide against just a single execution because prior work DR detectors, being not integrated with a MC, are not intended to uncover different results under subsequent runs.
				% Define a "stress tester" as a class of bug detectors where they are intended to [...]
				% Maybe say: Do there exist any stress-tester DR detectors where they are intended to produce new results under reruns?

One might also wonder: Why is it fair to compare the data race bugs \quicksand~finds (10 CPU-hours)
against the candidates found by a prior work's single execution (less than a minute)?
We argue this comparison is meaningful because prior work data-race tools, being not integrated with a MC,
are not intended to uncover different results under subsequent runs.
One could run a data-race tool repeatedly for 10 CPU-hours, but the advantage of stateless MC over stress testing is already well-understood.
% TODO TODO TODO get a big citation for teh above sentence.

@@

% TODO: Future work.
% Future work: Add a way to configure even smaller subsets (eg "only mutex_locks called from site X") for cases where mx_lock and mx_unlock alone are still too big. Count the number of kernels for which this was the case.
However, larger state spaces often result from finer-grained synchronization,
which can indicate a more complicated concurrent algorithm that would require more rigorous verification than a program with a single global lock.
(analogous to a developer % changing an implementation to be finer graind and then re running the same unit test)

### 2015-10-17 pldi-16/intro.tex
write about free-re-malloc and add figure 1

% blah blah trite opening sentence
%As parallelism becomes ever more important for achieving high performance in modern-day programs,
%so too do advanced concurrency testing techniques become important for verifying the correctness of those programs.
Concurrency bugs are notoriously hard to find and reproduce because they only appear in specific thread interleavings, which arise at random during normal program execution.
{\em Stateless model checking} \cite{verisoft} offers a method for finding such bugs,
or verifying their absence,
%by systematically executing a program along as many distinct interleavings as possible,
by forcing a program to execute each distinct interleaving,
capturing and controlling this nondeterminism in a finite state space.
Unfortunately, these state spaces explode exponentially in the size of the input program.
Reduction techniques such as Dynamic Partial Order Reduction (DPOR) \cite{dpor} and Maximal Causality Reduction (MCR) \cite{mcr} expand the limits of feasible test completion,
and search ordering strategies such as Iterative Context Bounding (ICB) \cite{chess} allow bugs to be found sooner in a given exploration should they exist.

% Can I even make a claim this broad to begin with?
However, all stateless model checkers to date are bound by a fixed set of {\em preemption points}: code locations that define the granularity at which threads interleave.
% TODO: You repeat the following 1.5 sentence in the related work section; maybe you can eliminate this redundancy to save space?
For example, \textsc{CHESS} \cite{chess} preempts only on synchronization operations and library calls, which can miss lock-free shared memory races.
It provides an additional data-race analysis to report any violations of this model;
however, data-race analyses are prone to report false positives and benign races which require annotations or imprecise heuristics to reduce \cite{racerx,tsan,datacollider}.
%
On the other hand, SPIN \cite{spin}
is able to preempt threads around any shared memory access. Such fine granularity would automatically check if each data race is a real bug, but makes full state space completion intractable for all but the most rudimentary test inputs.
%
This work shows how to avoid making that tradeoff decision in advance.

>>>>

% blah blah trite opening sentence
%As parallelism becomes ever more important for achieving high performance in modern-day programs,
%so too do advanced concurrency testing techniques become important for verifying the correctness of those programs.
Concurrency bugs are notoriously hard to find and reproduce because they only appear in specific thread interleavings, which arise at random during normal program execution.
{\em Stateless model checking} \cite{verisoft} offers a method for finding such bugs,
or verifying their absence,
%by systematically executing a program along as many distinct interleavings as possible,
by forcing a program to execute each distinct interleaving,
capturing and controlling this nondeterminism in a finite state space.
Unfortunately, these state spaces explode exponentially in the size of the input program.
Reduction techniques such as Dynamic Partial Order Reduction \cite{dpor} and Maximal Causality Reduction \cite{mcr} expand the limits of feasible test completion,
and search ordering strategies such as Iterative Context Bounding \cite{chess} allow bugs to be found sooner in a given exploration should they exist.

% Can I even make a claim this broad to begin with?
However, all stateless model checkers to date are bound by a fixed set of {\em preemption points}: code locations that define the granularity at which threads interleave.
% TODO: You repeat the following 1.5 sentence in the related work section; maybe you can eliminate this redundancy to save space?
For example, \textsc{CHESS} \cite{chess} preempts only on synchronization operations and library calls, which can miss lock-free shared memory races.
It provides an additional data-race analysis to report any violations of this model;
however, data-race analyses are prone to report false positives
%and benign races
which require annotations or imprecise heuristics to reduce \cite{racerx,tsan,datacollider}.
%
On the other hand, SPIN \cite{spin}
is able to preempt threads around any shared memory access. Such fine granularity would automatically check if each data race is a real bug, but makes full state space completion intractable for all but the most rudimentary test inputs.
%
This work shows how to avoid making that tradeoff decision in advance.

### 2015-10-17 pldi-16/related.tex
write about free-re-malloc and add figure 1

{\bf Preempting at instruction granularity} is a prerequisite for using data-race PPs. However, every such prior tool we know of has serious drawbacks.
% TODO: Be proud of them instead of being dismissive.
{\textsc SKI} \cite{ski} approaches kernel code by choosing in advance a random set of instruction offsets from the start of the test,
which is more similar to stress testing or fuzzing than to exhaustive state space exploration.
% TODO: Make sure you talk about your mutex experiment, or rephrase this.
SPIN \cite{spin} specializes in verifying synchronization primitive implementations such as RCU, which is very similar to our \mxtest~experiment.
However, it does not employ Iterative Deepening, and requires programs to be written in the PROMELA DSL.
%so cannot check implementations directly.

@@

\newcommand\portend{\textsc{Portend}}
\portend~\cite{portend} is the most closely related work we have found.
It performs a single-pass analysis as normal, then explores a limited state space to classify candidates in a taxonomy of likely severity.
Compared to us, \portend~can detect non-failing races which nevertheless cause suspiciously different program output,
while we depend on directly detecting failures.
It also uses symbolic execution to test input nondeterminism along with schedule nondeterminism,
while we explore the latter only.
However, \portend~does not explore alternate interleavings in advance of knowing data-race locations,
which is necessary to expose some bugs (\sect{\ref{sec:eval-falseneg}}).
It is also limited to the POSIX synchronization API, so cannot verify lock implementations such as we do with \mxtest.
Future work could combine the two approaches, using \landslide~to produce new traces for \portend~to classify, or using \portend's analysis to inform \quicksand's state space priorities.

>>>>

{\bf Preempting at instruction granularity} is a prerequisite for using data-race PPs.
However, the resulting state space explosion demands that any such tool either
choose a small subset of instructions to consider as PPs.
or be limited to very small test inputs.
%However, every such prior tool we know of has serious drawbacks.
{\textsc SKI} \cite{ski} approaches kernel code by choosing in advance a random set of instruction offsets from the start of the test,
which is more similar to stress testing or fuzzing than to exhaustive state space exploration.
SPIN \cite{spin} specializes in verifying synchronization primitive implementations such as RCU, which is very similar to our \mxtest~experiment.
However, it cannot practicelly test full-size programs, and requires code to be written in the PROMELA DSL.
%so cannot check implementations directly.

@@

\newcommand\portend{\textsc{Portend}}
\portend~\cite{portend} is the most closely related work we have found.
It performs a single-pass analysis as normal, then explores a limited state space to classify candidates in a taxonomy of likely severity.
Compared to us, \portend~can also detect non-failing races which nevertheless cause suspiciously different program output,
while we depend on directly detecting failures.
It uses symbolic execution to test input nondeterminism as well as schedule nondeterminism,
while we explore the latter only.
However, \portend~does not explore alternate interleavings in advance of knowing data-race locations,
which is necessary to expose some bugs (\sect{\ref{sec:eval-falseneg}}).
It also assumes the POSIX synchronization API, so cannot verify lock implementations such as we do with \mxtest.
Future work could combine the two approaches, using \landslide~to produce new traces for \portend~to classify, or using \portend's analysis to inform \quicksand's state space priorities.
